---
title: "Exam_2017"
author: "Jacintha Johnson"
date: "8/3/2017"
output: html_document
---

#Question 1
```{r Question1, echo=TRUE}
#Import libraries to tidy and manipulate the data
library(dplyr)
library(tidyr)
library(ggplot2)

#import the data file
pain1 <- read.csv("https://dl.dropboxusercontent.com/u/11805474/painblogr/biostats/exam-data-2017/question_1.csv")

#1a) Group and location (which are two separate variables) are stored in one column. Pain intensity readings are in three separate columns, these can be stacked into one column.

#tidy and manipulate the data
pain1 %>%
  separate(Group.and.location, c("group", "location"), sep= "_", remove=TRUE)%>%
 select(location, Baseline:Week.8)%>%
gather(pain1, pain, Baseline:Week.8) -> tidied
tidied

ggplot(tidied, aes(x=pain1, y=pain)) + geom_point(shape=1) + geom_smooth(method=lm, se=FALSE) + ylab(label="Pain Intensity") + xlab("Time") + ggtitle("Relationsip between pain intensity over 3 time intervals at each location")


```

#Question 2
##Hypothesis
####Ho:There is no relationship between religion and fertility in Swiss states
####HA: Fertility does depend on religion in Swiss states
####Data is quantitative and continuous so the statistical test to use will be a multiple linear regression model. 
##Assumptions
####There is a linear relationship between the outcome variable and the independent variables. 
####It assumes the variables are normally distributed. 
####It assumes the independetn variables are not highlighy correlated with each otehr. 
####It assumes homoscedasticity
```{r Question2, echo=TRUE}
#Import libraries to tidy and manipulate the data
library(dplyr)
library(tidyr)

#import the data file
religion <- read.csv("https://dl.dropboxusercontent.com/u/11805474/painblogr/biostats/exam-data-2017/question_2.csv")

#get a summary of key values
summary(religion)

#use a scatterplot to plot the data as we will use a multiple linear regression model
plot(x= religion$State, y= religion$Fertility, ylab="Fertility", main= "Relationship betweeen religion and fertility in the late 19th century French-speaking Swiss states", xlab="States")

#use a linear regression model as the statistical test
Q2test <- lm(Fertility~Catholic, data = religion)

#get the outcomes of the test
summary(Q2test)

#df= 45, p-value= 0.001029
#The p-value is less than 0.05, therefore we reject the null hypothesis
#One can conclude there is a relationship between religion and fertility in these 19th century Swiss states
```

#Question 3
##Hypothesis
####Ho: There is no difference in crop yield between the 3 groups
####HA: There is a difference in the crop yields between the 3 groups

####This data is quantitative and continuos
####So one will use an two-way ANOVA test 
##Assumptions
####An ANOVA assumes independence of cases, normality, homoscedasticity, groups have same sample size, variances of populaations are equal. 
```{r Question3, echo=TRUE}

#Import libraries to tidy and manipulate the data
library(dplyr)
library(tidyr)

#import the data file
treatment <- read.csv("https://dl.dropboxusercontent.com/u/11805474/painblogr/biostats/exam-data-2017/question_3.csv")


#get a summary of key values
summary(treatment)

#Plot the data using a boxplot
boxplot(treatment, main="Relationship between condition and crop yield.", xlab="Condition", ylab="Crop Yield")

#perform statistical test (ANOVA)
test <- aov(control~treatment_1*treatment_2, data=treatment)
test
summary(test)

#treatment_1: p-value: 0.1024 , df = 1
#Treatment_2: p-value: 0.0648 , df = 1
#interaction: p-value: 0.1391 , df = 1
#p-values are greater than 0.05 therefore we accept the null hypothesis. #one can conclude there is no difference in the crop yield between the 3 conditions

```

#Question 4
##Hypothesis
####Ho: The variables are independent
####HA: The variables are dependent
####This data is categorical with 2 or more categories so one will use a chi-square test (Fishers to get an exact p-value)
##Assumptions
####the samples have been drawn randomly and a directional hypothesis is assumed
```{r Question4, echo=TRUE}

#Import libraries to tidy and manipulate the data
library(dplyr)
library(tidyr)

#import the data file
heartattack <- read.csv("https://dl.dropboxusercontent.com/u/11805474/painblogr/biostats/exam-data-2017/question_4.csv")

#get a summary of key values
summary(heartattack)

#contingency table of the data
contab <- table(heartattack$Diet, heartattack$Condition)
contab

#Perform Statistical test
fishtest <- fisher.test(contab)
fishtest

#p-value = 0.0004174
#alternative hypothesis: two.sided
#The p-value is less than 0.05 therefore one rejects the null hypothesis
#one can conclude the variables are dependent
#the Mediterranean diet is superior to the AHA diet
```
